Viscous Transport in Eroding Porous Media

### 2019-05-24 docs/transport/erosion_transport.tex
small changes to methods

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Shear Stress}
To compute the rate of erosion, we must compute the shear stress of the
double-layer potential
\begin{align}
  \tau(\xx) = \left(\nabla \DDD[\eeta](\xx) + 
    \nabla^T \DDD[\eeta](\xx) \right) \nn \cdot \ss,
\end{align} where $\nn$ and $\ss$ are the unit normal and tangent
vectors at $\xx$, respectively.  Each term of the gradient of the
velocity field can be written in terms of Cauchy integrals.  For
example,
\begin{align}
  \pderiv{u_1}{x_2} &= \Re(iv'[\tau_1](x)) + 
      \Re(i v''[y \cdot \eta](x)) -
      x_1\Re(i v''[\eta_1](x))  -
      \Re(v'[\eta_2](x)) - x_2 \Re(iv''[\eta_2](x)) \\
    &= -\Im(v''[\tau_1](x)) - 
      \Im(v''[y \cdot \eta](x)) +
      x_1\Im(v''[\eta_1](x)) -
      \Re(v'[\eta_2](x)) + x_2 \Im(v''[\eta_2](x)).
\end{align}
Similarly, the other three components of $\nabla \DDD[\eeta]$ can be
calculated
\begin{align}
  \pderiv{u_1}{x_1} &= \Re (v'[\tau_1](x)) + 
    \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
    x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
  \pderiv{u_1}{x_2} &= - \Im (v'[\tau_1](x)) - 
    \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
    \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_1} &= \Re (v'[\tau_2](x)) - 
    \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
    x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_2} &= -\Im (v'[\tau_2](x)) - 
    \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
    \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{s:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We have shown that the velocity, shear stress, vorticity, and pressure
of the double-layer potential can all be written in terms of Cauchy
integrals and its first two derivatives.  Therefore, we require
quadrature methods to evaluate the expressions in
equation~\eqref{eqn:cauchy}.  Since the integrands are all smooth and
periodic, the Trapezoid rule has spectral accuracy.  However, the
integrand is nearly singular if $x$ is close to $\bd\Omega$, and the
quadrature error is arbitrarily large if the resolution $N$ is fixed.
There have been a variety of near-singular integration schemes
developed, and we extend a method first outlined by Barnett et
al.~\cite{bar-wu-vee2015} which we call the {\em Barycentric quadrature
rule}.  This method first uses the density function $\tau(y)$ to compute
the boundary data of $v$, denoted by $v^-(y)$.  Then, it makes use
of the identity
\begin{align}
  v(x) = \int_{\bd\Omega} \frac{v^{-}(y)}{y-x} \, dy, 
  \quad x \in \Omega.
\end{align}

@@

At each GMRES iteration, we must compute the velocity at each
discretization point on $\bd\Omega$.  To reduce the cost of computing
the velocity, we start by using the trapezoid rule~\eqref{eqn:trap}
accelerated with the fast multipole method (FMM)~\cite{gre-rok1987},
which requires $\mathcal{O}(N)$ operations, and we call this velocity
$v_\trap(\xx)$.  For target points that are sufficiently far from
$\bd\Omega$, $v_\trap(\xx)$ is sufficiently accurate.  However, this
velocity is inaccurate at points that are too close to $\bd\Omega$, and
the velocity must be corrected.  Assuming that $\xx$ is too close to
$\gamma_k$, only the velocity due to $\gamma_k$ must be corrected. We
first subtract the velocity contribution of $\vv(\xx)$ due to
$\gamma_k$.  Then, the Barycentric quadrature rule is used to compute
the velocity due to $\gamma_i$ with more accuracy.  That is, the
final velocity is
\begin{align}
  \vv(\xx) = \vv_\trap(\xx) - \frac{L_k}{N_k} \sum_{j=1}^{N_\iin} 
    D(\xx,\yy^k_j) \eeta(\yy_j) + \vv^k_\bary(\xx),
\end{align}
where $v^k_\bary(\xx)$ is the velocity at $\xx$ resulting from applying
the Barycentric quadrature rule due to $\gamma_k$.  This method
naturally extends to points that are close to $\Gamma$, and to the
scenario when $\xx$ is simultaneously close to multiple components of
$\bd\Omega$.

We do not use any fast summation methods to apply the Barycentric
quadrature rule.  Therefore, to it is imperative that we only apply
the Barycentric quadrature at target points where they are required.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{s:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We do some simulations.  The parameters are
\begin{itemize}
  \item time step size
  \item regularization and smoothing parameters
  \item resolution on bodies and solid wall
\end{itemize}
{\color{red} In this section, we present the numerical simulations of bodies eroding in the fluid. The computational domain $\Omega$ is a rectangle $[-3,3]\times[-1,1]$ with rounded off corners and we place bodies only in $[-1,1]\times[-1,1]$. We impose the fixed-pressure-drop condition on the simulation
 smoothing parameters of $\eps$ and $\sigma $.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A single body close to a wall}
\todo[inline]{Do a single body that is extremely close one of the two
horizontal walls.  This can be compared to some of the
results~\cite{mit-spa2017}.  Do for a variety of distances to
demonstrate how close we can put it to the wall with no need to increase
the resolution}
\begin{figure}[htbp]
\begin{center}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r1h_shear}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r0d5h_shear}
\includegraphics[width = 0.35 \textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{A single body eroding in Stokes flow. Color indicates the shear stress on the body. In this simulation, we use three minimum body-to-wall initial distances $d=h$ (left), $h/2$ (middle), and $h/10$ (right) with $N_\iin = 256$ discretization points on the body, $N_\out = 1024$ points on the outer boundary, a time-step of $\Dt=0.0001$, and smoothing parameters of $\eps$ and $\sigma $.}
\label{} 
\end{center}
\end{figure}
{\color{red} We first consider a single circular body in a shear flow. 
The domain $\Omega$ is a rectangle $[-3,3]\times[-1,1]$ with rounded corners. 
The velocity of flow from the left side of $\Omega$ is $v(x_1,x_2)=x_2+1$. 
Let the initial radius of the body be $r$, 
the initial length of the body be $L$, $N_\iin$ and 
$N_\out$ be the number of discretization points 
on the body and wall respectively, 
$h=L/N_\iin$ be the resolution mesh size of the body, 
and $d$ be the minimum initial distance from body to wall. 
To investigate the shape of eroding body when it is close to wall, 
we set $r=0.4$, $N_\iin=256$, and $N_\out=1024$ 
with respect to three different $d=h, h/2$, and $ h/10$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Cauchy Integral Representation of the Shear Stress}
To compute the rate of erosion, we must compute the shear stress of the
double-layer potential
\begin{align}
  \tau(\xx) = \left(\nabla \DDD[\eeta](\xx) + 
    \nabla^T \DDD[\eeta](\xx) \right) \nn \cdot \ss,
\end{align} where $\nn$ and $\ss$ are the unit normal and tangent
vectors at $\xx$, respectively.  Each term of the gradient of the
velocity field can be written in terms of Cauchy integrals.  For
example,
\begin{align}
  \pderiv{u_1}{x_2} &= \Re(iv'[\tau_1](x)) + 
      \Re(i v''[y \cdot \eta](x)) -
      x_1\Re(i v''[\eta_1](x))  -
      \Re(v'[\eta_2](x)) - x_2 \Re(iv''[\eta_2](x)) \\
    &= -\Im(v''[\tau_1](x)) - 
      \Im(v''[y \cdot \eta](x)) +
      x_1\Im(v''[\eta_1](x)) -
      \Re(v'[\eta_2](x)) + x_2 \Im(v''[\eta_2](x)).
\end{align}
Similarly, the other three components of $\nabla \DDD[\eeta]$ can be
calculated
\begin{align}
  \pderiv{u_1}{x_1} &= +\Re (v'[\tau_1](x)) + 
    \Re (v''[y\cdot\eta](x)) - \Re (v'[\eta_1](x)) - 
    x_1\Re (v''[\eta_1](x)) - x_2\Re (v''[\eta_2](x)), \\
  \pderiv{u_1}{x_2} &= - \Im (v'[\tau_1](x)) - 
    \Im (v''[y\cdot\eta](x)) + x_1\Im (v''[\eta_1](x)) - 
    \Re (v'[\eta_2](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_1} &= +\Re (v'[\tau_2](x)) - 
    \Im (v''[y\cdot\eta](x)) + \Im (v'[\eta_1](x)) +
    x_1\Im (v''[\eta_1](x)) + x_2\Im (v''[\eta_2](x)), \\
  \pderiv{u_2}{x_2} &= -\Im (v'[\tau_2](x)) - 
    \Re (v''[y\cdot\eta](x)) + x_1\Re (v''[\eta_1](x)) +
    \Im (v'[\eta_2](x)) + x_2\Re (v''[\eta_2](x)).
\end{align}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Methods}
\label{s:method}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We have shown that the velocity, shear stress, vorticity, and pressure
of the double-layer potential can all be written in terms of Cauchy
integrals and its first two derivatives.  Therefore, we require
quadrature methods to evaluate the expressions in
equation~\eqref{eqn:cauchy}.  Since the integrands are all smooth and
periodic, the Trapezoid rule has spectral accuracy.  However, the
integrand is nearly singular if $x$ is close to $\bd\Omega$, and the
quadrature error is arbitrarily large for a fixed resolution.  There
have been a variety of near-singular integration schemes developed, and
we extend a method first outlined by Barnett et
al.~\cite{bar-wu-vee2015} which we call the {\em Barycentric quadrature
rule}.  This method first uses the density function $\tau(y)$ to compute
the boundary data of $v$, denoted by $v^-(y)$.  Then, it makes use of
the identity
\begin{align}
  v(x) = \int_{\bd\Omega} \frac{v^{-}(y)}{y-x} \, dy, 
  \quad x \in \Omega.
\end{align}

@@

At each GMRES iteration, we must compute the velocity at each
discretization point on $\bd\Omega$.  To reduce the cost of computing
the velocity, we start by using the trapezoid rule~\eqref{eqn:trap}
accelerated with the fast multipole method (FMM)~\cite{gre-rok1987},
which requires $\mathcal{O}(N)$ operations, and we call this velocity
$\vv_\trap(\xx)$.  For target points that are sufficiently far from
$\bd\Omega$, $\vv_\trap(\xx)$ is sufficiently accurate.  However, this
velocity is inaccurate at points that are too close to $\bd\Omega$, and
the velocity must be corrected.  Assuming that $\xx$ is too close to
$\gamma_k$, only the velocity due to $\gamma_k$ must be corrected. We
first subtract the velocity contribution of $\vv(\xx)$ due to
$\gamma_k$.  Then, the Barycentric quadrature rule is used to compute
the velocity due to $\gamma_i$ with more accuracy.  That is, the final
velocity is
\begin{align}
  \vv(\xx) = \vv_\trap(\xx) - \frac{L_k}{N_k} \sum_{j=1}^{N_\iin} 
    D(\xx,\yy^k_j) \eeta(\yy_j) + \vv^k_\bary(\xx),
\end{align}
where $\vv^k_\bary(\xx)$ is the velocity at $\xx$ resulting from
applying the Barycentric quadrature rule due to $\gamma_k$.  This method
naturally extends to points that are close to $\Gamma$, and to the
scenario when $\xx$ is simultaneously close to multiple components of
$\bd\Omega$.

Similarly, to compute the vorticity and shear stress, the trapezoid rule
is first used.  Then, if $\xx$ is close to $\gamma_k$, the trapezoid
quadrature rule contribution due to $\gamma_k$ is subtracted and the
Barycentric quadrature rule contribution due to $\gamma_k$ is added.
Since the shear stress and vorticity are only computed once per time
step, we use a direct summation rather than the FMM.  Relative to the
cost of applying GMRES with the FMM, the additional once-per-time-step
cost is minimal.

We also do not use any fast summation methods to apply the Barycentric
quadrature rule.  Since this quadrature rule is applied at every GMRES
iteration, it is imperative that we only apply the Barycentric
quadrature at target points where they are required.  As a rule of
thumb, the trapezoid rule due to $\gamma_k$ is accurate if
\begin{align}
  d(\xx,\gamma_k) = \inf_{\yy \in \gamma_k} \|\xx - \yy\| < 
    5 \frac{L_k}{N_\iin}.
\end{align}
To avoid having to check if a target point $\xx$ satisifies

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical Results}
\label{s:results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We do some simulations.  The parameters are
\begin{itemize}
  \item time step size
  \item regularization and smoothing parameters
  \item resolution on bodies and solid wall
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A single body close to a wall}
\todo[inline]{Do a single body that is extremely close one of the two
horizontal walls.  This can be compared to some of the
results~\cite{mit-spa2017}.  Do for a variety of distances to
demonstrate how close we can put it to the wall with no need to increase
the resolution}
\begin{figure}[htbp]
\begin{center}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r1h_shear}
\includegraphics[width = 0.30 \textwidth]{./figs/1b_0d4r0d5h_shear}
\includegraphics[width = 0.35 \textwidth]{./figs/1b_0d4r0d1h_shear}
\caption{A single body eroding in Stokes flow. Color indicates the shear stress on the body. In this simulation, we use three minimum body-to-wall initial distances $d=h$(left), $h/2$(middle), and $h/10$(right) with $N_\iin = 256$ discretization points on the body, $N_\out = 1024$ points on the outer boundary, a time-step of $\Dt$, and smoothing parameters of $\eps $ and $\sigma $.}
\label{} 
\end{center}
\end{figure}
{\color{red} We first consider a single circular body in a shear flow. 
The domain $\Omega$ is a rectangle $[-3,3]\times[-1,1]$ with rounded corners. 
The velocity of flow from the left side of $\Omega$ is $v(x_1,x_2)=x_2+1$. 
Let the initial radius of the body be $r$, 
the initial length of the body be $L$, $N_\iin$ and 
$N_\out$ be the number of discretization points 
on the body and wall respectively, 
$h=L/N_\iin$ be the resolution mesh size of the body, 
and $d$ be the minimum initial distance from body to wall. 
To investigate the shape of eroding body when it is close to wall, 
we set $r=0.4$, $N_\iin=256$, and $N_\out=1024$ 
with respect to three different $d=h, h/2$, and $ h/10$.
